Value Ordering, Discrepancies, and Restarts in Subgraph Solvers: \\ Some Half-Baked Mucking Around

### 2017-11-09 paper/paper.tex
Bits

We could try to introduce tiebreaking mechanisms to further refine this heuristic. However, given
the extreme cost of making an incorrect branching choice early on in search, we believe it is more
fruitful to investigate alternatives to simple backtracking (depth-first) search. This paper
evaluates two such approaches: discrepancy search (which perform poorly), and restarts (which are
extremely beneficial). To make restarts work, we introduce a new value-ordering heuristic, which is
nearly the same as ``highest degree first'' except with small amounts of heavily biased randomness.
The effects of the changes we propose are shown in \cref{figure:scatter-by-family}: we can make a
state-of-the-art algorithm perform much better on a large number of satisfiable instances, without
substantially degrading its performance on unsatisfiable instances.

@@

    \caption{The effects of using depth-bounded discrepancy search.}\label{figure:dds}
\end{figure}

@@

?? \cref{figure:dds}. Sometimes extremely beneficial on satisfiable instances. However, on both
unsatisfiable and satisfiable instances, the overheads are often large, and the algorithm is not
better in aggregate (even when only considering satisfiable instances).

@@

\begin{figure}[p]
    \centering
    \includegraphics*{gen-graph-scatter-random.pdf}
    \caption{Instance by instance comparison of the basic algorithm with the degree heuristic,
    versus random + restarts.}
    \label{figure:scatter-random}
\end{figure}

@@

    \caption{Instance by instance comparison of the basic algorithm with the degree heuristic,
    versus degree-biased + restarts.}
    \label{figure:scatter-by-family}
\end{figure}

>>>>

We could try to introduce tiebreaking mechanisms to further refine this heuristic. However, given
the extreme cost of making an incorrect branching choice early on in search, we believe it is more
fruitful to investigate alternatives to simple backtracking (depth-first) search. This paper
evaluates two such approaches: discrepancy search (which perform poorly), and restarts (which are
extremely beneficial). To make restarts work, we introduce a new value-ordering heuristic, which is
nearly the same as ``highest degree first'' except with small amounts of heavily biased randomness.
The effects of the changes we propose are shown in \cref{figure:scatter-by-family}: we can make a
state-of-the-art algorithm perform much better on a large number of satisfiable instances, whilst
performing worse only rarely on satisfiable instances, and never on unsatisfiable instances.
Further, these properties are \emph{not} shared by discrepancy searches (\cref{figure:scatter-dds})
or by random value-ordering with restarts (\cref{figure:scatter-random}).

@@

    \caption{The effects of using depth-bounded discrepancy search.}\label{figure:scatter-dds}
\end{figure}

@@

?? \cref{figure:scatter-dds}. Sometimes extremely beneficial on satisfiable instances. However, on both
unsatisfiable and satisfiable instances, the overheads are often large, and the algorithm is not
better in aggregate (even when only considering satisfiable instances).

@@

\begin{figure}[p]
    \centering
    \includegraphics*{gen-graph-scatter-random.pdf}
    \caption{Instance by instance comparison of the basic algorithm with the degree heuristic,
    versus random value-ordering with restarts.}
    \label{figure:scatter-random}
\end{figure}

@@

    \caption{Instance by instance comparison of the basic algorithm with the degree heuristic,
    versus degree-biased value ordering with restarts.}
    \label{figure:scatter-by-family}
\end{figure}
